Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-17 paper/agree.tex
prep for pull on paper

\newsavebox{\sw}
\begin{lrbox}{\sw}
\begin{lstlisting}[style=agree,numbers=left]
eq req : bool = event(AutomationRequest);
eq avl : bool = event(AirVehicleLocation);
eq wp : bool = event(Waypoint);
eq strt: bool = event(Start);
eq alrt : bool = event(Alert);

assume "Automation requests are well-formed" :
  req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
assume "Air vehicle locations are well-formed" :
  avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);    
assume "One automation request in flight at a time" :
  true -> 
  (req => pre(Historically(not req) or Since(not req, strt)));

guarantee "Waypoints coincide with air vehicle locations":
  wp => avl;
guarantee "Starts include a new waypoint" :
  strt => wp;
guarantee "Waypoints are well-formed" : 
  wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Starts within one cycle of requests if not alerting" :
  (strt => ((not alrt) and req)) -> 
  (strt => ((not alrt) and (req or pre(req))));
guarantee "Alert if not started within one cycle of requests" :
    true -> ((pre(req and not strt) and not strt) => alrt);
guarantee "Once alerted always alerted" :
  not alrt or (Once(alrt) and alrt);
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \scalebox{0.62}{\usebox{\sw}}
  \end{center}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

@@

\newsavebox{\flt}
\begin{lrbox}{\flt}
\begin{lstlisting}[style=agree,numbers=left]
eq policy : bool = 
  WELL_FORMED_AUTOMATION_RESPONSE(Input);

guarantee Filter_Output
  "Filter output is well-formed" :
  if event(Input) and policy then 
    event(Output) and Output = Input
  else not event(Output);
\end{lstlisting}
\end{lrbox}

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree,numbers=left]
const is_latched : bool = true;
const MAX_LATENCY : int = 1;

eq rsp : bool = event(Response);
eq req : bool = event(Request);

eq isPending : bool = Since(not rsp, req and not rsp);
eq latency : int = 0 -> (if req then 0 else pre(latency) + 1);

eq policy : bool = (rsp => req) ->
                   (    (isPending => latency < MAX_LATENCY)   
                    and (rsp => (req or pre(isPending))));
eq alert : bool = (not policy) -> 
                  ((is_latched and pre(alert)) or not policy);

assume "One outstanding request at a time" :
  (true -> (req => not pre(isPending)));

guarantee "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee "Output if not alerted" :
  if (not(alert) and rsp) then
    event(Output) and (Output = Response)
  else
    not (event(Output));    
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\flt}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance filter.}
  \label{fig:filter}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\mntr}} \\
    \end{tabular}
  \end{center}
  \caption{Contract specification for high-assurance monitor.}
  \label{fig:monitor}
\end{figure}

@@

The AGREE contract specification generated by the transform is shown in \figref{fig:filter}.
The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
The resulting AGREE specification for the monitor in this example is shown in \figref{fig:monitor}.
The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
This behavior is one of the several options available in the dialogue.
The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.

>>>>

\newsavebox{\sw}
\begin{lrbox}{\sw}
\begin{lstlisting}[style=agree]
eq req : bool = event(AutomationRequest);
eq avl : bool = event(AirVehicleLocation);
eq wp : bool = event(Waypoint);
eq rsp: bool = event(Start);
eq alrt : bool = event(Alert);

assume "Automation request is well-formed" :
    req => WELL_FORMED_AUTOMATION_REQUEST(AutomationRequest);
assume "Air vehicle location is well-formed" :
    avl => WELL_FORMED_WAYPOINT(AirVehicleLocation);

eq current : bool = (req = rsp);
eq previous : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = current or previous;
eq since : bool =  alrt or (alrt and (false -> pre(since)));

guarantee "Start includes a waypoint" :
    rsp => wp;
guarantee "Locations required after the start waypoint" :
    (wp and not rsp) => avl;
guarantee "Waypoint is well-formed" :
    wp => WELL_FORMED_WAYPOINT(Waypoint);
guarantee "Alert if start is not bounded relative to a request" :
    policy or since;
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \scalebox{0.60}{\usebox{\sw}}
  \end{center}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

@@

\newsavebox{\flt}
\begin{lrbox}{\flt}
\begin{lstlisting}[style=agree]
eq policy : bool =
  WELL_FORMED_AUTOMATION_RESPONSE(Input);
guarantee Filter_Output "Filter output is well-formed" :
  if event(Input) and policy then
    event(Output) and Output = Input
  else not event(Output);
\end{lstlisting}
\end{lrbox}

\newsavebox{\mntr}
\begin{lrbox}{\mntr}
\begin{lstlisting}[style=agree]
const is_latched : bool =
  Get_Property(this, CASE_Properties::Monitor_Latched);
eq rsp : bool = event(Response);
eq req : bool = event(Request);
eq current : bool = (req = rsp);
eq previous : bool = (req and not rsp) ->
                      pre(req and not rsp) and (not req and rsp);
eq policy : bool = current or previous;
eq alert : bool = (not policy)
                -> ((is_latched and pre(alert)) or not policy);
guarantee Monitor_Alert
  "Alert port tracks alert variable" :
  event(Alert) = alert;
guarantee Monitor_Output
  "Output if not alerted" :
  if event(Alert) then (not event(Output)) else
  if event(Response) then (event(Output) and (Output = Response))
  else (not event(Output));
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\flt}} \\
      (a) \\
      \scalebox{0.60}{\usebox{\mntr}} \\
      (b)
    \end{tabular}
  \end{center}
  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
  \label{fig:assurance}
\end{figure}

@@

The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a).
The guarantee is stylized for synthesis and completely defines the meaning of the output under every possible input.
The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b).
The \texttt{is\_latched} value makes the alert persistent, meaning that once the alert is raised, it is always raised.
This behavior is one of the several options available in the dialogue.
The definition for \texttt{policy} is taken by the system developer from the contract in \figref{fig:sw}.
As before, the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.
